### 2020-01-26 paper/sex_selection_spacing-ver3.tex
Combine questions and results in Introduction

India has changed significantly over the last three decades:
the economy has grown substantially,
education levels have increased for both males and females,
and the total fertility rate has fallen to the point where, at 2.2, it is just
above replacement 
\citep{Bosworth2008,Guilmoto2013,Dharmalingam2014,
International-Institute-for-Population-Sciences-IIPS2017}.

@@

Finally, birth spacing has long served as a measure of son preference, with strong
son preference associated with shorter birth intervals, the fewer sons a family has 
\citep{ben-porath76b,Leung1988}.
Now, longer birth intervals in the absence of sons, rather than an indicator of 
lower son preference, may instead arise because couples with a strong preference 
for sons use sex selection.

@@

First, how did average birth intervals change over time for different groups and how do 
these changes relate to the use of sex selection?
The hypothesis is that in the absence of sex selection son preference leads to shorter 
spacing the fewer boys a family has, whereas son preference increases spacing the fewer 
boys a family has when prenatal sex selection is available, possibly to the point where
we observe a reversal of the traditional spacing pattern.
Families with son preference who---for one reason or another---do not use prenatal sex 
selection likely continue to have shorter birth spacing.
Based on the prior literature on sex selection, I focus on how birth intervals have
changed for different education levels, urban or rural, and the sex composition of 
prior children.

Second, how has the distribution of birth spacing changed?
Although examining average birth intervals helps us understand the overall changes in 
spacing behavior, it tells us little about how these changes came about.
For example, the motivations that lead to an increase in average spacing are different if 
the increase comes mainly from parents trying to avoid very short birth intervals or from 
continued use of sex selection to ensure a son.

Third, to what extent does TFR reflect cohort fertility?
An advantage of the empirical model is that I can predict fertility behavior taking into
account both the likelihoods of parity progression and having a child of a specific sex.
I compare those predictions with fertility rate estimates based on the same data to
show how the increases in birth spacing has led to a substantial overestimate of the
decline in fertility [TK too imprecise? underestimate of the level, rather?]

@@

There are four main results.
First, the intervals between births have increased for all education groups over the four 
decades.
These increases are larger the higher the parity and the higher the education 
level.
Second, those women who are most likely to use sex selection---well-educated women with no
sons---have seen the largest increases in birth intervals.
As a result, well-educated women now show a reversal of the traditional spacing 
pattern, with the longest birth intervals for women with no sons.
Third, women who are least likely to use sex selection, those with no education in
rural areas, still follow the standard pattern of short spacing when they have girls and 
little evidence of the use of sex selection.
Finally, the traditional pattern of longer average spacing with sons appear to arise from 
more very long intervals, rather than from fewer very short birth intervals; across 
education groups, there appears to be little difference in the likelihood of having a very 
short birth interval across sex compositions, especially for lower parities.

>>>>

India has changed significantly over the last four decades:
the economy has grown substantially,
education levels have increased for both males and females,
and the total fertility rate has fallen to the point where, at 2.2, it is just
above replacement 
\citep{Bosworth2008,Guilmoto2013,Dharmalingam2014,
International-Institute-for-Population-Sciences-IIPS2017}.

@@

Finally, birth spacing has long served as a measure of son preference, with strong
son preference associated with shorter birth intervals, the fewer sons a family has 
\citep{ben-porath76b,Leung1988}.
Now, longer birth intervals in the absence of sons, rather than an indicator of 
lower son preference, may instead arise because couples with a strong preference 
for sons use sex selection, possibly to the point where we observe a reversal of 
the traditional spacing pattern.
Families with son preference who---for one reason or another---do not use prenatal 
sex selection likely continue to have shorter birth spacing.

@@

First, how did average birth intervals change over time for different groups, and how do 
these changes relate to the use of sex selection?
To show how sex selection has affected birth spacing, I focus on how birth intervals have
changed for different education levels, urban or rural, and the sex composition of 
prior children.

Birth intervals have increased for all education groups over the four decades, 
with larger increases, the higher the parity and the higher the education level.
Women who are most likely to use sex selection---well-educated women with no
sons---have seen the most substantial increases in birth intervals.
[TK check this!] As a result, well-educated women now show a reversal of the traditional spacing 
pattern, with the longest intervals for women with no sons.
Those least likely to use sex selection, women with no education in rural areas, 
still follow the traditional pattern of short spacing when they have girls and 
little evidence of the use of sex selection.

Second, how has the distribution of birth spacing changed?
Although examiniong average birth intervals helps us understand the overall changes in 
spacing behavior, it tells us little about how these changes came about.
For example, the motivations that lead to an increase in average spacing are different if 
the increase comes mainly from parents trying to avoid very short birth intervals or from 
continued use of sex selection to ensure a son.

[TK this is unclear]
The traditional pattern of longer average spacing with sons appears to arise from 
an increase in the number of very long intervals, rather than from fewer very short 
birth intervals. 
Across education groups, there appears to be little difference in the likelihood of 
having a very short birth interval across sex compositions, especially for lower parities.

Third, to what extent does TFR reflect cohort fertility?
An advantage of the empirical model is that I can predict fertility behavior taking into
account both the likelihood of parity progression and having a child of a specific sex.
I compare those predictions with fertility rate estimates based on the same data to
show how the increases in birth spacing have led to a substantial overestimate of how 
fast fertility has declined. [TK too imprecise? underestimate of the level, rather?]

@@

[TK results]
